# canadaHCD
Access Canadian Historical Climate Data from R. The Government of Canada's Historical Climate Data [website](http://climate.weather.gc.ca/index_e.html) provides access to hourly, daily, and monthly weather records for stations throughout Canada.

These are raw data that have undergone some quality control, but issues such as changes in station location are unmanged; the data for the original `stationID` stops at a certain point and a new `stationID` continues recording. For a more curated data set for climate change research at broad spatial and temporal scales see the [Adjusted and Homogenized Canadian Climate Data (AHCCD)](http://ec.gc.ca/dccha-ahccd/default.asp?lang=En&n=B1F8423A-1).

#### Released version
[![CRAN version](http://www.r-pkg.org/badges/version/canadaHCD)](http://cran.rstudio.com/web/packages/canadaHCD/index.html) [![](http://cranlogs.r-pkg.org/badges/grand-total/canadaHCD)](http://cran.rstudio.com/web/packages/canadaHCD/index.html)

#### Build status
[![Build Status](https://travis-ci.org/gavinsimpson/canadaHCD.svg?branch=master)](https://travis-ci.org/gavinsimpson/canadaHCD)  [![Build status](https://ci.appveyor.com/api/projects/status/e3ptg9drviavci71/branch/master?svg=true)](https://ci.appveyor.com/project/gavinsimpson/canadahcd/branch/master)  [![codecov.io](https://codecov.io/github/gavinsimpson/canadaHCD/coverage.svg?branch=master)](https://codecov.io/github/gavinsimpson/canadaHCD?branch=master)

## Installation

*canadaHCD* is still under active development towards a 0.1 release. In the meantime, if you wish to use the package, please install it from this github repo, which is most easily achieved using Hadley Wickham's *devtools* package:

```{r, eval = FALSE}
## install.packages("devtools")
devtools::install_github("gavinsimpson/canadaHCD")
```

## Usage

Say I'm interested in climate data for stations in Yellowknife, I can search for all known `stationID`s with `"Yellowknife"` in their name using `find_station()`

```{r find-station}
library("canadaHCD")
find_station("Yellowknife")
```

To download the monthly HCD from `YELLOWKNIFE HYDRO` I can use `hcd_monthly()`, providing it with the `StationID` for that particular weather station

```{r download-monthly, results = "hide"}
yh <- hcd_monthly(1707)
```

The data are returned as a [*tibble*](https://cran.r-project.org/web/packages/tibble/vignettes/tibble.html) (a `tbl_df`), which shows a compact version of the data frame.

```{r show-data}
yh
```

You should be able to work with these objects mostly as if they were data frames.

Allthough not yet exposed through any functions in the package, you can access a snapshot of the station metadata via the `canadaHCD:::station_data` data frame.

```{r station-data}
canadaHCD:::station_data
```

If we wanted to know which resolutions of data were available for the `YELLOWKNIFE HYDRO` station, we can extract certain columns from the station data object

```{r extract-station_data}
id <- grep("YELLOWKNIFE HYDRO", canadaHCD:::station_data$Name)
vars <- c("HourlyFirstYr", "HourlyLastYr", "DailyFirstYr", "DailyLastYr", "MonthlyFirstYr", "MonthlyLastYr")
canadaHCD:::station_data[id, vars]
```
